Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-11-02 03_marlin_paper.qmd
pre-pub edits

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.1"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

@@

bio-economic modeling
conservation planning
fisheries
marine protected areas
marine spatial planning
spatio-temporal modeling

@@

Communities around the world are increasingly looking to spatial strategies for
managing marine ecosystems . For example, the "30x30" movement calls for
implementation of marine protected areas (MPAs), a form of spatial management,
across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
references therein]. Recent agreements such as the United Nations (UN) agreement
on the conservation and sustainable use of marine biodiversity of areas beyond
national jurisdiction (BBNJ) calls for expanded area-based management in the
high seas. However, designing spatial management strategies to achieve desired
objectives -- which may include recovery and resilience of overfished species,
increased food production and economic well being, and equitable distribution of
benefits - is not straightforward. Different species have different resilience
to fishing, distributions in space and time, and value to fishing fleets;
fishing fleets themselves can have varying ranges of incentives and reactions to
policy structures. These dynamics pose challenges even for single-species
assessment and management strategies , which are only amplified when we consider
management policies specifically designed around multiple species, fleets, and
spatial features in the oceans [@field2006a].

@@

The front end of the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] to facilitate use, while the underlying population model is
written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for
speed. Users on standard computers should be able to simulate one age-structured
population distributed across two-dimensional spatial surface represented by a
ten by ten grid of cells over 20 years in fractions of a second.

@@

Below we provide a summary of the `marlin` model, as well as details of the case
studies. We note that `marlin` is a complex software package with many options
to control detailed behaviors. Given that these specific features will evolve
over time, we do not describe in detail every option of the model here. Rather,
we focus on the equations and methods describing the core foundations of our
tool for efficiently simulating bio-economic dynamics in aquatic systems.

@@

`marlin` simulates the dynamics of one or more species, currently best
representing fish-like species, using age-structured population dynamics. Ages
are then converted to lengths using the von Bertalanffy growth equation with
log-normally distributed variation in the length at age. Each time step, species
move throughout the simulated seascape using both diffusion and "taxis" (active
movement towards preferred habitat), experience natural and potentially fishing
mortality, and if in a spawning season spawn using one of the possible forms of
density dependence implemented in the model, with the potential to include
auto-correlated stochastic deviations in the amount of offspring produced,
generally called "recruitment deviates".

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse-seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack (*Katsuwonus pelamis*) tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
It is important to note that contact selectivity reflects the ability of the
fishing method in question to capture fish of different sizes that come into
"contact" with the gear. The contact selectivity of each of the specified
*metiers* will then interact with the distribution of fish sizes and fishing
effort in space simulated by `marlin` to produce a net "population" selectivity,
which may differ from the individual contact selectivities of each of the
*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
fishing effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* section), for example in proportion to
profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

The dynamics of each fleet can be modified by management in a variety of ways.
Managers can impose size limits for individual species within each fleet. They
can also impose total catch quotas for each species in each time step. When
quotas are activated, if the total catch across all fleets for a given species
under the model's effort dynamics would exceed the allowable quota for that
species, the total effort for each fleet is decreased proportional to its
contribution to the total catch until the quota is satisfied. As an alternative
to catch quotas, managers can set an effort cap per fleet, which prevents effort
from exceeding a given amount under open access dynamics, though effort may be
reduced below this cap if required by the profit equation (i.e. the fleet can
choose to fish less than the quota or effort cap).

`marlin` also allows users to specify closed fishing seasons for one or more
species in the system. When all species in a simulation share the same closed
seasons, no fishing occurs in those seasons. When not all species share the same
closed seasons, fishing still occurs in the closed seasons, but the fleet avoids
catching the species for which fishing is closed in a given season, assuming no
discard mortality. When closed seasons are in effect, the amount of effort in
the next open season is a function of the effort and profits in the last open
fishing season. Lastly, managers can specify locations of no-take MPAs, which
can change in size and location if needed. When MPAs are implemented, fishing
effort that used to operate inside the MPAs can either leave the fishery, or be
redistributed outside the MPA (the default behavior).

Users can use the model to explore general properties of spatially explicit
social-ecological systems, or can parameterize it to explore key uncertainties
of more tactical applications. It is very important note that the parameters of
this model cannot be fit to data directly within the `marlin` package; users
must set model parameters themselves based on externally available data and
their best judgement.

On the biological side, at a minimum users must for each species they wish to
simulate supply the common or scientific name of the species in question, a
measure of the level of fishing intensity the species is experiencing at the
start of the simulation, and the diffusion rates for adult and larvae. Given a
common or scientific name, the model will then input default life history
parameters based on FishLife [internet connection required, @thorson2020],
though users are encouraged to check these values and supply their own life
history parameter values when possible.

@@

The degree of fishing pressure exerted by a given set of fishing fleets on each
species is a function of a range of parameters including the total amount of
effort (*E*), the contact selectivity ogives ($\alpha$,the curve specifying the
proportion of individuals at a given age or size that will be caught by a unit
of fishing mortality), fishing cost (*C*), the spatial distribution of the
species affected by the fleet, the relative prices across species ($\Pi$), and
the catchability coefficient of each fleet for each species ($q$). `marlin`
provides a tuning option to help users achieve desired biological outcomes from
their fleets.

@@

Users can also tune the fleet dynamics by specifying the ratio of costs to
revenues. Price data for use in the model can be obtained relatively simply
through literature reviews, market surveys, or local experts, so long as they
are converted to the same units of biomass used in the model. However, cost
parameters are more complicated, as translating say cost per day of fishing into
the same representation of fishing effort used in the model is not straight
forward to accomplish. As an alternative, users can specify an equilibrium
cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet
in question, and `marlin` will tune the cost parameter to achieve this desired
cost to revenue ratio given the other parameters in the model (e.g. price and
biomass of the target species). So for example, while users may not have access
to exact cost parameters, they may have a general sense that when operating
under ideal conditions a given fleet has a profit margin of 10%, i.e. the cost
to revenue ratio is 90%. They can input the cost to revenue ratio as 0.9 then
and allow `marlin` to tune the cost parameters accordingly.

@@

Each fishing fleet decides how to allocate its effort in space based on one of
four possible spatial allocation strategies. The ideal free distribution (IFD)
is the standard method for distributing fishing fleets in spaces (see
@gillis2003 and references therein). However, analytical solutions to the IFD
present a number of complications for our model. Our model is capable of
tracking multiple fleets at once. This means that the IFD for a single fleet
would commonly be modeled as a Nash-equilibrium based on the actions of each of
the individual fleet conditional on the actions of all other fleets. While
possible to implement, this would slow down our model runs to the point of
practically preventing large-scale evaluation of the spatial management
policies.

As such we explored a series of "next best" fleet distribution algorithms. While
not the IFD in any individual time step, over time they start to approximate the
IFD, as each assumes that fleets base their decisions for the current time step
on the outcomes in the prior timestep, meaning that the impacts of the actions
of other fishing fleets are somewhat internalized. The timeline required for the
fleets to reach an equilibrium distribution will vary and depend on factors such
as the population dynamics of the species in question (e.g. movement, habitat
heterogeneity, growth rates, density dependence) and the number and degree of
competition across fleets. Users should explore different simulation times to
ensure that any results they wish to use are not simply a reflection of the
fleet dynamics fluctuating on their way to an equilibrium condition.

@@

The underlying population model used is an age structured single-species model
in the manner of @ovando2021a. The population model requires many parameters.
However, if the user supplies either a scientific or common name for the species
in question (scientific preferred), the model will supply default values for
that species based on the values reported in FishLife [internet connection
required, @thorson2020]. FishLife provides estimates of core life history
parameters for fish species based on a model integrating published values and
evolutionary connections. FishLife provides reasonable default values for given
species, but these default values should be not be taken as definitive and users
should input best available parameters specific to the stock in question if they
wish to best represent the dynamics of their specific system, and should check
the plausibility of the default values prior to use. All of the life history
values can be replaced with user-supplied values as needed and should be for
more tactical applications; the default values simply ensure that reasonable and
internally consistent values for the species in question are supplied as a
starting point.

@@

For both the diffusion and taxis matrices, we allow for the inclusion of
physical barriers to movement (i.e. land). Pairs of patches that are adjacent
but in which one or both patches are a barrier to movement are set as
non-adjacent. The CTMC model then produces movement dynamics that move around
barriers rather than over them. This is different than setting habitat in a
patch to zero; an animal will not preferentially move towards a patch with zero
habitat all else being equal, but can move over a patch with zero habitat
towards another better patch. An example might be a species that lives and feeds
on coastal habitats transiting through a patch of open ocean to reach a new
feeding ground. In contrast, a physical barrier represents something like a
peninsula that an animal must move around to transit from one side to another.

@@

Fisheries management is often concerned with measuring stock status relative to
maximum sustainable yield (MSY) based reference points, though the exact level
of stock status relative to MSY reference points desired by management or
societal objectives may vary widely. MSY based reference points present a
problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of
fishery selectivity. Fishery selectivity in this model can vary by fleet, and
species can be distributed unevenly in both space and time. This means that the
net effective fishing selectivity on a species can vary depending on the
dynamics at a given moment, making the definition of an equilibrium concept such
as MSY challenging [@berger2017 and references therein].

As such, we do not report MSY based reference points in the model by default,
with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
tends to be better conserved across selectivity patterns as it is primarily a
function of the mortality and flow of recruits relative to the maturity ogive.
There are many different strategies for estimating reference points in spatially
explicit systems [@kapur2021a]. We leave it to users to define and find relevant
reference point as required by their specific needs.

@@

Each of the case studies contains too many parameters and options to be
succinctly presented in the text here, and we are concerned that attempting to
duplicate all of the parameters of the model within the text would simply
introduce errors and confusion. As such, and since these results are for
illustrative purposes only, readers should consult the accompanying code to view
the precise details of each simulation. Targeted applications must carefully
consider and document all decisions made around model parameters.

@@

To our knowledge, MPAs are never randomly placed; their locations are decided
based on stakeholder negotiation given available information and objectives. In
theory, the design of MPA networks can be optimized through the use of a
modeling framework, and depending on the validity of the model, this process is
likely to produce better outcomes than manually-designed strategies
[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
becomes increasingly difficult as the range of objectives and the complexity of
the model increase. Therefore, we focus here on the design and performance of these more rule of thumb design
strategies that may be more accessible to a wider range of users. We allow all
MPAs to be designed in a mosaic fashion in these examples, but users can easily
extend the analysis to compare outcomes between contiguous (MPA is made up of
one continuously connected block) and mosaic (MPAs can be separated in space)
MPA designs [@pons2022] by developing and passing their own MPA design
algorithms to the model.

@@

In our coastal coral reef example, we model the dynamics of four tropical
Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
(*Lutjanus malabaricus*), a deep-reef snapper (*Pristipomoides filamentosus*),
and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
a mass migration to a spawning aggregation once per year, followed by the
sharks. Shallow-reef snappers stay in reefs closer to shore above a steep
drop-off year-round, while deep-water snappers stay in the deeper reefs past the
drop-off (@fig-coral).

These species are targeted by two different fleets. Fleet One primarily targets
the grouper and near-shore snapper populations, but will land any incidentally
captured sharks. Fleet One has a logistic selectivity pattern for all species,
as they retain any fish caught for consumption or sale. Fleet One is totally
dependent on fishing for their livelihood, meaning the local community takes
advantage of every possible opportunity to fish, and as such we model it as a
"constant effort" fishery. Due to having less efficient boats, Fleet One has a
higher cost per distance coefficient than Fleet Two. Fleet One's home port is
located on near the site of the grouper spawning grounds.

@@

We used `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e. the
selected MPA design can be as patchwork as dictated by the data). We ran the
simulation for in quarter year time steps for 20 years ($\Delta_t = 1/4$) and
set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
patches for a total area of 2,000 KM^2^.

@@

We model our pelagic case study loosely on the characteristics of the Western
and Central Pacific Ocean (WCPO) tuna fisheries. Note that this is an illustrative example only and simulated stock status, species distributions, or outcome projections presented here should not be interpreted as a indicative of the current or future state of the WCPO. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
purse-seine fishery that primarily targets skipjack tunas for bulk canning.
Contact selectivities were modeled as logistic for the longline fleet, and
dome-shaped for the purse-seine fleet, with selectivity values set to broadly
match the patterns reported in relevant [stock
assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
for each species. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of the
selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
range shift. Specifically, we simulate an extreme example where the centroid of
each population shifts northward at a rate of 0.2 patches per year over a 20
year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
in the starting year, and then held that network constant over the years of the
experiment, running one simulation with and another without the climate-drive
range shift. We then compared the effects of this range shift on food production
and conservation outcomes from MPA networks designed based on the pre-range
shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
patches each with a side length of roughly 311 KM, for a total area of
14,000,000 KM^2^, broadly commensurate with the area of the WCPO.

@@

MPAs were capable of producing a range of positive and negative outcomes for
food security and conservation in the coral reef case study depending on the
design strategy used and the size of MPA implemented. Both of the MPA design
strategies were capable of increasing fisheries yield for the more "industrial"
Fleet Two, up to a value of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
even when MPAs covered more than 50% of the simulated area. However, the more
"artisanal" Fleet One only benefited from MPAs under the *Spawning Ground*
design strategy, with a maximum increase of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
the *Target Fishing* design strategy produced a roughly linear decrease in
fishing yields as a function of increasing MPA size (@fig-coral-results A).

On the conservation front, MPAs were uniformly beneficial to the spawning
biomass of all species under the *Target Fishing* design strategy, with the most
rapid increases in spawning biomass for the deep snapper population. The
*Spawning Ground* strategy primarily benefited the shallow snapper population,
producing little change in the grouper population and decreasing spawning
biomass of both the deep snapper and reef shark populations for MPA sizes
covering less than 50% of the simulated area (@fig-coral-results B).

@@

Marine ecosystems are driven by complex social-ecological dynamics. Communities
must often make decisions on how to manage these systems based on limited
empirical evidence. Modeling tools such as the one presented here can help users
answer scientific questions and design marine management policies informed by a
better understanding of sensitivities to key uncertainties. In this paper we
have documented the core system of equations underlying `marlin`, and
demonstrated how `marlin` can be used to support marine resource management.

@@

The variation in both the magnitude and distribution of costs and benefits
stemming from alternative design choices and states of nature shown in our case
studies demonstrates the value in our method. For example, our coastal coral
reef example illustrates both the potential for MPAs to benefit conservation and
food production in these systems, and the potential for the same MPA to benefit
some fleets and species while harming others when fishing fleets affected by an
MPA do not share the same objectives and species are not uniformly distributed.
Our pelagic case study illustrates how `marlin` can be used to assess the
climate resiliency of spatial management strategies, and identify strategies
that best meet the needs of both current and future conditions. Results from
this type of work could be used to help prioritize communities and species at
particular risk to climate change impacts.

Our coastal coral reef case study results show the sensitivity of conservation
and food security outcomes of MPAs to alternative design scenarios, holding
constant other social-ecological variables. The *Spawning Ground* MPA network
was able to provide more equitable yield outcomes across the two fishing
communities, whereas the *Target Fishing* MPA network only benefited Fleet Two (@fig-coral-results A). This is because Fleet One has two
primary fishing grounds; the spawning grounds, and the offshore area where the
deep snapper live. Fleet Two primarily fishes in the northern portions of the
simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
losses fishing grounds on the spawning grounds, it can compensate for this by
moving offshore and fishing the deep snapper population harder, while also being
able to fish some of the spillover of snapper from the spawning ground closure.
This displacement of Fleet One's fishing effort is why biomass of deep snapper
actually declines under the *Spawning Ground* strategy.

>>>>

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx:
    reference-doc: template.docx
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.1"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

@@

Bio-economic modeling; Conservation planning; Fisheries; Marine protected areas; Marine spatial planning; Spatio-temporal modeling

@@

Communities around the world are increasingly looking to spatial strategies for
managing marine ecosystems. For example, the "30x30" movement calls for
implementation of marine protected areas (MPAs), a form of spatial management,
across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
references therein]. Recent agreements such as the United Nations (UN) agreement
on the conservation and sustainable use of marine biodiversity of areas beyond
national jurisdiction (BBNJ) calls for expanded area-based management in the
high seas. However, designing spatial management strategies to achieve desired
objectives -- which may include recovery and resilience of overfished species,
increased food production and economic well being, and equitable distribution of
benefits - is not straightforward. Different species have different resilience
to fishing, distributions in space and time, and value to fishing fleets;
fishing fleets themselves can have varying ranges of incentives and reactions to
policy structures. These dynamics pose challenges even for single-species
assessment and management strategies, which are only amplified when we consider
management policies designed around multiple species, fleets, and
spatial features in the oceans [@field2006a].

@@

The front end of the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] to facilitate use, while the underlying population model is
written in C++, integrated through the `Rcpp` package [@eddelbuettel2018]. Users on standard computers should be able to simulate one age-structured population distributed across two-dimensional spatial surface represented by a
ten by ten grid of cells over 20 years in fractions of a second. It is very important to note that the parameters of
this model cannot be fit to data directly within the `marlin` package; users
must set model parameters themselves based on externally available data and
their best judgement.

@@

Below we provide a summary of the `marlin` model, as well as details of the case
studies.

@@

`marlin` simulates the dynamics of one or more species, currently best
representing fish-like species, using age-structured population dynamics. Ages
are then converted to lengths using the von Bertalanffy growth equation with
log-normally distributed variation in the length at age. Each time step, species
move throughout the simulated seascape using both diffusion and "taxis" (active
movement towards preferred habitat), experience natural and potentially fishing
mortality, and potentially spawn using one of the possible forms of
density dependence implemented in the model, with the ability to include
auto-correlated stochastic deviations in the amount of offspring produced,
generally called "recruitment deviates".

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *métier*). For example, both a longline and purse-seine fleet may
capture bigeye (*Thunnus obesus*, Scombridae) and skipjack (*Katsuwonus pelamis*, Scombridae) tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
It is important to note that contact selectivity reflects the ability of the
fishing method in question to capture fish of different sizes that come into
"contact" with the gear. The contact selectivity of each of the specified
*métiers* will then interact with the distribution of fish sizes and fishing
effort in space simulated by `marlin` to produce a net "population" selectivity,
which may differ from the individual contact selectivities of each of the
*métiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
fishing effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* section), for example in proportion to
profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

The dynamics of each fleet can be modified by management in a variety of ways.
Managers can impose size limits for individual species within each fleet. They
can also impose total catch quotas for each species in each time step. When
quotas are activated, if the total catch across all fleets for a given species
under the model's effort dynamics would exceed the allowable quota for that
species, the total effort for each fleet is decreased proportional to its
contribution to the total catch until the quota is satisfied. As an alternative
to catch quotas, managers can set an effort cap per fleet, which prevents effort
from exceeding a given amount under open access dynamics, though effort may be
reduced below this cap if required by the profit equation (i.e. the fleet can
choose to fish less than the quota or effort cap). `marlin` also allows users to specify closed fishing seasons for one or more
species in the system. Lastly, managers can specify locations of no-take MPAs, which
can change in size and location if needed. When MPAs are implemented, fishing
effort that used to operate inside the MPAs can either leave the fishery, or be
redistributed outside the MPA (the default behavior).

On the biological side, at a minimum users must for each species being simulated supply the common or scientific name of the species in question, a
measure of the level of fishing intensity the species is experiencing at the
start of the simulation, and the diffusion rates for adult and larvae. Given a
common or scientific name, the model will then input default life history
parameters based on FishLife [internet connection required, @thorson2020],
though users are encouraged to check these values and supply their own life
history parameter values when possible.

@@

The degree of fishing pressure exerted by a given set of fishing fleets on each
species is a function of a range of parameters including the total amount of
effort (*E*), the contact selectivity ogives ($\alpha$), fishing cost (*C*), the spatial distribution of the
species affected by the fleet, the relative prices across species ($\Pi$), and
the catchability coefficient of each fleet for each species ($q$). `marlin`
provides a tuning option to help users achieve desired biological outcomes from
their fleets.

@@

Users can also tune the fleet dynamics by specifying the ratio of costs to
revenues. Price data for use in the model can be obtained relatively simply
through literature reviews, market surveys, or local experts. However, cost
parameters are more complicated, as translating say cost per day of fishing into
the same representation of fishing effort used in the model is not straight
forward to accomplish. As an alternative, users can specify an equilibrium
cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet
in question, and `marlin` will tune the cost parameter to achieve this desired
cost to revenue ratio given the other parameters in the model.

@@

Each fishing fleet decides how to allocate its effort in space based on one of
four possible spatial allocation strategies. The ideal free distribution (IFD)
is the standard method for distributing fishing fleets in spaces (see
@gillis2003 and references therein). However, analytical solutions to the IFD
present a number of complications for our model. The IFD for a single fleet
would commonly be modeled as a Nash-equilibrium based on the actions of each of
the individual fleet conditional on the actions of all other fleets. While
possible to implement, this would slow down our model runs to the point of
practically preventing large-scale evaluation of the spatial management
policies.

As such we explored a series of "next best" fleet distribution algorithms. While
not the IFD in any individual time step, over time they start to approximate the
IFD, as each assumes that fleets base their decisions for the current time step
on the outcomes in the prior timestep, meaning that the impacts of the actions
of other fishing fleets are eventually accounted for. The timeline required for the
fleets to reach an equilibrium distribution will vary and depend on factors such
as the population dynamics of the species in question and the number and degree of
competition across fleets. Users should explore different simulation times to
ensure that any results they wish to use are not simply a reflection of the
fleet dynamics fluctuating on their way to an equilibrium condition.

@@

The underlying population model used is an age structured single-species model
in the manner of @ovando2021a. The population model requires many parameters.
However, if the user supplies either a scientific or common name for the species
in question (scientific preferred), the model will supply default values for
that species based on the values reported in FishLife [@thorson2020]. FishLife provides estimates of core life history
parameters for fish species based on a model integrating published values and
evolutionary connections. FishLife provides reasonable default values for given
species, but these default values should be not be taken as definitive and users
should check the default values and input best available parameters specific to the stock in question if they
wish to best represent the dynamics of their specific system.

@@

For both the diffusion and taxis matrices, we allow for the inclusion of
physical barriers to movement (i.e. land). Pairs of patches that are adjacent
but in which one or both patches are a barrier to movement are set as
non-adjacent. The CTMC model then produces movement dynamics that move around
barriers rather than over them.

@@

Fisheries management is often concerned with measuring stock status relative to
maximum sustainable yield (MSY) based reference points, though the exact level
of stock status relative to MSY reference points desired by societal objectives may vary widely. MSY based reference points present a
problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of
fishery selectivity. Fishery selectivity in this model can vary by fleet, and
species can be distributed unevenly in both space and time. This means that the
net effective fishing selectivity on a species can vary depending on the
dynamics at a given moment, making the definition of an equilibrium concept such
as MSY challenging [@berger2017 and references therein].

As such, we do not report MSY based reference points in the model by default.
There are many different strategies for estimating reference points in spatially
explicit systems [@kapur2021a]. We leave it to users to define and find relevant
reference points as required by their specific needs.

@@

Each of the case studies contains too many parameters and options to be
succinctly presented in the text here. Readers should consult the accompanying code to view
the precise details of each simulation. Targeted applications must carefully
consider and document all decisions made around model parameters.

@@

In theory, the design of MPA networks can be optimized through the use of a
modeling framework, and depending on the validity of the model, this process is
likely to produce better outcomes than manually-designed strategies
[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
becomes increasingly difficult as the range of objectives and the complexity of
the model increase. Therefore, we focus here on the design and performance of these more rule of thumb design
strategies that may be more accessible to a wider range of users. We allow all
MPAs to be designed in a mosaic fashion in these examples, but users can easily
extend the analysis to compare outcomes between contiguous (MPA is made up of
one continuously connected block) and mosaic (MPAs can be separated in space)
MPA designs [@pons2022].

@@

In our coastal coral reef example, we model the dynamics of four tropical
Pacific species: a grouper (*Epinephelus fuscoguttatus*, Serranidae), a shallow-reef snapper
(*Lutjanus malabaricus*, Lutjanidae), a deep-reef snapper (*Pristipomoides filamentosus*, Lutjanidae),
and a reef shark (*Carcharhinus amblyrhynchos*, Carcharhinidae). The simulated groupers undergo
a mass migration to a spawning aggregation once per year, followed by the
sharks. Shallow-reef snappers stay in reefs closer to shore above a steep
drop-off year-round, while deep-water snappers stay in the deeper reefs past the
drop-off (@fig-coral).

These species are targeted by two different fleets. Fleet One primarily targets
the grouper and near-shore snapper populations, but will land any incidentally
captured sharks. Fleet One has a logistic selectivity pattern for all species,
as they retain any fish caught for consumption or sale. Fleet One is totally
dependent on fishing for their livelihood, meaning the local community takes
advantage of every possible opportunity to fish, and as such we model it as a
"constant effort" fishery. Due to having less efficient boats, Fleet One has a
higher cost per distance coefficient than Fleet Two. Fleet One's home port is
located near the site of the grouper spawning grounds.

@@

We used `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity. We ran the
simulation for in quarter year time steps for 20 years ($\Delta_t = 1/4$) and
set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
patches for a total area of 2,000 KM^2^.

@@

We model our pelagic case study loosely on the characteristics of the Western
and Central Pacific Ocean (WCPO) tuna fisheries. Note that this is an illustrative example only and simulated stock status, species distributions,  and projections presented here should not be interpreted as a indicative of the current or future state of the WCPO. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*, Carcharhinidae) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin (*Thunnus albacares*, Scombridae) for high-grade consumption, and a
purse-seine fishery that primarily targets skipjack tunas for bulk canning.
Contact selectivities were modeled as logistic for the longline fleet, and
dome-shaped for the purse-seine fleet. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of the
selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
range shift. Specifically, we simulate an extreme example where the centroid of
each population shifts northward at a rate of ~62km per year over a 20
year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
in the starting year, and then held that network constant over the years of the
experiment, running one simulation with and another without the climate-drive
range shift. We then compared the effects of this range shift on food production
and conservation outcomes from MPA networks designed based on the pre-range
shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
patches each with a side length of roughly 311 KM, for a total area of
14e6 KM^2^, broadly commensurate with the area of the WCPO.

@@

MPAs were capable of producing a range of positive and negative outcomes for
food security and conservation in the coral reef case study depending on the
design strategy used and the size of MPA implemented. Both of the MPA design
strategies were capable of increasing fisheries yield for
Fleet Two, up to a value of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
even when MPAs covered more than 50% of the simulated area. However Fleet One only benefited from MPAs under the *Spawning Ground*
design strategy, with a maximum increase of
`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
the *Target Fishing* design strategy produced a roughly linear decrease in
fishing yields as a function of increasing MPA size (@fig-coral-results A).

MPAs were uniformly beneficial to the spawning
biomass of all species under the *Target Fishing* design strategy, with the most
rapid increases in spawning biomass for the deep snapper population. The
*Spawning Ground* strategy primarily benefited the shallow snapper population,
producing little change in the grouper population and decreasing spawning
biomass of both the deep snapper and reef shark populations for MPA sizes
covering less than 50% of the simulated area (@fig-coral-results B).

@@

Marine ecosystems are driven by complex social-ecological dynamics. Communities
must often make decisions on how to manage these systems based on limited
empirical evidence. Modeling tools such as the one presented here can help users
answer scientific questions and design marine management policies informed by a
better understanding of sensitivities to key uncertainties.

@@

Our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and
food production in these systems, and the potential for the same MPA to benefit
some fleets and species while harming others when fishing fleets affected by an
MPA do not share the same objectives and species are not uniformly distributed.
Our pelagic case study illustrates how `marlin` can be used to assess the
climate resiliency of spatial management strategies, and identify strategies
that best meet the needs of both current and future conditions. Results from
this type of work could be used to help prioritize communities and species at
particular risk to climate change impacts.

Holding constant other social-ecological variables in the coral reef example, the *Spawning Ground* MPA network
was able to provide more equitable yield outcomes across the two fishing
communities, whereas the *Target Fishing* MPA network only benefited Fleet Two (@fig-coral-results A). This is because Fleet One has two
primary fishing grounds; the spawning grounds, and the offshore area where the
deep snapper live. Fleet Two primarily fishes in the northern portions of the
simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
losses fishing grounds on the spawning grounds, it can compensate for this by
moving offshore and fishing the deep snapper population harder, while also being
able to fish some of the spillover of snapper from the spawning ground closure.
This displacement of Fleet One's fishing effort is why biomass of deep snapper
actually declines under the *Spawning Ground* strategy.

Conversely, most of the catch in the coral reef fishery comes from the nearshore and deep
water snapper populations, which overlaps with both of Fleet One's fishing
grounds. The *Target Fishing* strategy then begins closing both the nearshore
and offshore snapper fishing grounds, resulting in too much of a loss in fishing
grounds for Fleet One to be offset by spillover from the MPAs. As a result of
primarily placing MPAs on Fleet One's fishing grounds, Fleet Two gains spillover
benefits at little cost to their fishing grounds until the MPAs reach their
fishing grounds once protection nears 100%.
